\chapter{Introduction}\label{ch:intro}
The modern world is moving towards digitalization at a high pace, which aims to make it smart and interconnected. 
This tranformation comes with the ever increasing use of connected devices and services, which oftentimes carry sensitive data and are not always protected sufficiently.
As a result, there is a rising number of cyber crime incidents and subsequent damages estimated to more than \$1 trillion worldwide.

Compoundig on top of the situation is a mismatch between demand and supply of Cybersecurity professionals and a general public, which is in large parts not aware of the risks posed by the use of connected IT technology.
Cybercriminals can often easily exploit these unaware people to gain access to resticted systems and data.
%People who are unaware of potential risks are often a weak spot and can be exploited by cyber criminals to access systems and networks. 
To combat this problem, there's a need for training Cybersecurity professionals, as well as to raise awareness in every level of education and the general public.
As a method for raising awareness, teaching priciples of Cybersecurity and the methods to apply them, live demonstration and hands-on learning with realistic risk scenarios can be an effective tool. \cite{mariano2024wifi}\cite{Cybersec_Edu}


\section{Goal}
An effective method to ensure the security of a system is pentesting\footnote{penetration testing, see \cref{ch:basics_cybersec}}, which helps to discover weaknesses in soft- and hardware.

The goal of this project is to develop a platform with a flexible and expandible architecture, which can be used for demonstration and practice of pentesting in Cybersecurity education.
A Raspberry PI will be used as the main device to create a platform, which is portable and easy to use.
It should be able to reliably set up testing environments for various fields of application, to allow a lecturer to demonstrate cyberattacks and students to practice penetration testing in a realistic enviroment.
In recent years, trends like the Internet of Things (IoT) have produced many headless\footnote{without any means for direct interaction} devices, which are often linked to smartphones for user interaction, using WiFi or Bluetooth for communication.
For this reason, wireless communication security in the form of WiFi technology will be implemented as the first application specific testing environment.
To teach secure design of web applications, the device should be able to host a server with the OWASP Juice Shop, which again ties into the security of IoT devices, because they are oftentimes accessed via web interfaces, which are often the source of vulnerabilities\cite[page~174]{Hellmann_2023}.


\section{Overview}
\begin{itemize}
    \item \cref{ch:basics_cybersec} explains the principles of it security and the difference between Cybersecurity and it security.
    \item \cref{ch:basics_wifi} explores the history of wifi technology and explains its architecture, function and security mechanisms.
    \item \cref{ch:documentation} explains the architecture of the hard- and software of the RCSL.
    \item \cref{ch:development} documents the development process and the challenges faced along the process.
    \item \cref{ch:testing} explains the penetration tests conducted and documents the results of testing them on the RCSL.
    \item \cref{ch:edu_use} outlines the challenges of Cybersecurity education and discusses usecases for the RCSL in the field.
\end{itemize}