\chapter{Basics}\label{ch:basics}

\section{Cybersecurity}
Through the progression of digitalization, Cybersecurity is of ever increasing relevance to the security of IT-Systems.
An increasing number of devices are connected to the internet, making them part of the cyberspace and therefore potentially vulnerable to cyberattacks.
Almost all sectors of our lives are at risk, when our homes, infrastructure, media devices, industries and governments are connected to the internet.

Cybersecurity aims to protect these devices by achieving the well-known CIA triad, which is an acronym for Confidentiality, Integrity, and Availability. \cite[chapter~Attacking Integrity Controls]{Oriyano_2017}
    



\section{WiFi History}

In 1985 the american Federal Communications Commission ruled serveral bands in the 2.4 GHz radio band to be unlicensed, meaning everyone can broadcast radiosignals in these frequencies.
At the beginning of the 90's, this lead to the development of different wireless communication protocols by multiple companies.
To ensure interoperability between these protocols, the IEEE 802.11 standard was created and released, and WLAN was born.
In 1999 the Wi-Fi Alliance (WFA) was founded to further improve compatibility, nowadays consisting of over 750 international companies. \cite[page~1-3]{Sankaran_Gulasekaran_2021}, \cite{Wikipedia_2025}

Since its inception over 25 years ago, Wi-Fi has gone through constant development, leading to six major generations of the technology.
Improvments in speed, efficiency and security have lead to the technology becoming a "fundamental utility [...], that everyone [...] expects to be available everywhere" \cite[page~1]{Sankaran_Gulasekaran_2021}.
WiFi connects billions of devices, like computers, smartphones, game consoles or IoT devices to the internet, generating a value of around 4.9 trillion USD worldwide \cite[page~1]{Sankaran_Gulasekaran_2021}.

WiFi operates at the physical (PHY) and the medium access control (MAC) layer of the OSI model, sending ethernet frames over radio waves in the 2.4 and 5 Ghz frequency band. 
The Mac layer is responsible for seting rules on tranmission and reception of data, connection management and power management.
Data sent on the MAC layer then gets converted to radio waveforms by the PHY layer, which is also responsible for converting received data back to the MAC protocol. \cite[page~4-7]{Sankaran_Gulasekaran_2021}

\section{MAC layer}

\subsection{Architecture}

Commonly WiFi networks have access points (AP), which are connected to the internet via ethernet, and stations (STA), which are provided with internet access over WiFi by the AP.
WiFi has a MAC layer, which is designed to be flexible and can be configured in the following topologies:
\begin{itemize}
    \item \textbf{BSS}: Basic Service Set
    \item \textbf{IBSS}: ad hoc / peer-to-peer / Independent BSS
    \item \textbf{WDS}: Wireless Distribution System
    \item \textbf{MBSS}: Mesh BSS
\end{itemize}

%evtl. Grafik einfügen

BSS describes the most common and basic configuration, where multiple STAs are connected to one AP, which has wired connection to the internet.
IBSS networks dont have any APs, instead the STAs are connected directly to each other.
WDS can be used, when the reach of an AP needs to be extended, but there is no wired connection available. 
A secondary AP can be installed which receives backhaul connection from another AP.
Mesh Networks are complex BSS networks with multiple APs, which the STAs can hop between.
This type of network can improve the signal coverage and resilience and is often used in companies or public institutions with many STAs. \cite[page~8-9]{Sankaran_Gulasekaran_2021}

\subsection{MAC Frames}

When data is sent over the MAC layer, it gets divided into chunks, which are called MAC Frames.
MAC Frames have three components: the header, the body and the frame check sequence.
The header contains information about the source and destination, the sequence and serveral control functions.
In a BSS network it has the following functions:

\begin{itemize}
    \item \textbf{Frame Control} contains 16 Bits which serve different control functions
        \subitem \textbf{protocol version} describes which version of the MAC protocol this frame uses
        \subitem \textbf{type} describes the type of MAC frame is beeing used: there are data, management, control and acknowledgment frames
        \subitem \textbf{subtype} describes the subtype of the frame, e.g. association request or beacon frame
        \subitem \textbf{from/to ds} indicates the direction of travel
        \subitem \textbf{more fragment} indicates more following fragments
        \subitem \textbf{retry} indicates that the current frame is a retransmission
        \subitem \textbf{power management} gives information about the power state of the STA
        \subitem \textbf{more data} indicates more frames are to be transmitted 
        \subitem \textbf{protected frame} indicates whether a frame is encrypted
        \subitem \textbf{order} set to 1 if the order of processing is important
    \item \textbf{duration} indicates the time of the transmission, therefore how long the medium is occupied
    \item \textbf{adress 1} contains the receiver address (RA), which can be the MAC address of the destination STA or the BSSID of the AP, depending on the direction, the frame travels
    \item \textbf{adress 2} contains the transmitter adress (TA), which can be the BSSID or the source address.
    \item \textbf{adress 3} contains either the source or destination address
    \item \textbf{sequence control}
        \subitem \textbf{fragment number} ensures the correct processing sequence of the fragments
        \subitem \textbf{sequence number} number of the frame
\end{itemize} 

The Frame body then contains the payload, meaning the actual information to be sent.
Every type of MAC frame is configured a different way to suit their specific function.
"While data frames are meant to carry data, management frames are used for connection management and control frames [...] for triggering specific actions.
Control frames have a smaller MAC header and no payload data [...].
Management frames have a structured payload composed of some fixed length fields and one or more variable length fields called information elements (IEs)" \cite[page~12]{Sankaran_Gulasekaran_2021}

After the frame body comes the frame control sequence (FCS), which ensures the integrity of the frame.
The FCS is a cyclic redundancy check (CRC), which is number calculated from the frame header and body content.
When the frame is compromised in traffic, the result of the CRC changes, therefore the error can be identified. \cite{GeeksforGeeks_2023} \cite[page~10-12]{Sankaran_Gulasekaran_2021}

\subsection{Medium Access}
In LAN and WLAN networks all devices share the same medium, therefore only one device at a time can transmit information.
Ethernet networks use full-duplex transceivers, meaning a device can transmit and receive data at the same time.
This allows the transceiver to transmit and simlutaniously listen for collisions with other frames.
Collision detection is quite effective, but can not be used in WiFi networks, because WiFi transceivers are only capable of half-duplex tranmission.
WiFi instead uses a medium access control referred to as carrier sense multiple access collision avoidance (CSMA-CA).

When using this technique, the medium gets blocked during transmission and is freed by an acknowledgment frame, sent by the receiving device.
All STAs start off in the random backoff state, where they listen for traffic.
After the random backoff counter has expired and no traffic is present, the STA will start transmission.
CSMA-CA can be described by the way "humans converse in a group, where every individual would first listen to check if someone else is talking, wait for the person talking to finish, and then start talking after a brief random pause" \cite[page~12]{Sankaran_Gulasekaran_2021}. \cite[chapter~Attacking Integrity Controls]{Oriyano_2017} \cite[page~12-17]{Sankaran_Gulasekaran_2021}

\subsection{Connection Establishment and Termination}
Before a connection can be established, the STA first needs to discover the AP it wants to connect to.
This can be done by either active or passive scanning of all available channels.

During active scanning, the STA transmitts a management frames, called probe requests on all channels to all SSIDs in the vicinity.
The APs answer with a probe respose frame, which contains the AP's SSID, BSSID and its capabilities.

Passive scanning works by the STA listening for beacon frames.
Beacon frames are periodically sent out by APs and also contain information on SSID, BSSID and capability. \cite[page~18]{Sankaran_Gulasekaran_2021}


\subsection{PHY layer}
The first generations of WiFi used transmit in the 2.4 Ghz band, but due to it beeing used by many other devices, such as cordless phones, computer peripherals, bluetooth devices or microwave ovens, WiFi 4 started using the 5 Ghz band, in order to improve connectivity.

\subsection{Security}
